COMP 6721: Project Report

### 2019-03-27 report/paper.tex
Add report.

\begin{document}
%
\title{COMP 6721: Project Report}
\author{Martin Khannouz}
%
\institute{40093517 \email{martin.khannouz@gmail.com}}
%
\maketitle              % typeset the header of the contribution
%
\section*{Introduction}
This project report describes how the first
project of \textit{Introduction to Artificial
Intelligence} was developed. The first section
covers the environment and the tools used for the
project. The heuristic section describes each
heuristic and how they were improved. The
difficulties section summerizes the difficulties
encountered during the project and how they were
overcome.

@@

\section{Heuristics}
Five heuristics were developed in this project
even though only two were made for competitive
use. One heuristic is the naive function for the
second deliverable. Another is a random heuristic
that systematicaly return a new random number. It
was used to spot runtime errors in \textit{MinMax} and
\textit{MonteCarlo} algorithms. The \textit{Basic}
heuristics only check for winning situation and
was used for basic controls.
The two other heuristics are described in
subsection of their own.

@@

\clearpage
\subsection{Vspace}
The goal of the Vspace heuristic is to give value to empty
cells in addition to counting rows of similar
symbols. Similarly to the previous heuristic,
Vspace count the size of the lines, but for every
line connected to an empty cell, it uses the
size of this line as a value for the empty cell.

@@

				\draw [->,thick] (A) -- (B);
		\end{tikzpicture}
		\caption{Example of the Vspace process for the
		color player based on the board in
		Figure~\ref{fig:board-example}. the grey cells
		represent cells that contain nothing
		interesting. On the right board, green cells
		are spaces available for the next turn and
		purple cells are spaces that need at least two
		cards to be filled.}
		\label{fig:color-vspace}
\end{figure}
\clearpage
\section{Difficulties}
\begin{itemize}
		\item Slowness of python, (numpy first)
		\item Version changing between Python 3.7.2 and 3.5.1 (serialization of regex)
\end{itemize}
\section{Analyze}
\begin{itemize}
		\item How is the pruning depth 3 with a
				recycling moves: no prun = 26s, prun=0.3s.
		\item How are the performance at the tournament.
\end{itemize}

>>>>

\begin{document}
%
\title{COMP 6721: Project Report}
\author{Martin Khannouz}
%
\institute{40093517 \email{martin.khannouz@gmail.com}}
%
\maketitle              % typeset the header of the contribution
%
\section*{Introduction}
This project report describes how the first
project of \textit{Introduction to Artificial
Intelligence} was developed. The first section
covers the environment and the tools used for the
project. The heuristic section describes each
%TODO Briefly describe what a heuristic is.
heuristic and how they were improved. The
difficulties section summerizes the difficulties
encountered during the project and how they were
overcome.

@@

\subsection{Tweaks in \textit{MinMax}}
The original \textit{MinMax} algorithm explore the
complete tree of possibilities and applies the
heuristic function on the leaf. However, the
heuristic functions described below do not take in
account who won first. For instance, a sequence of
move could be prefered if a player scores three
line of four even if his opponent already won at
the previous move with only one line of four.

To prevent this behavior, the MinMax algorithm was
tweaked to check for winning condition after
playing each possible move at any depth. This
modification prevent the exploration of game
states already won by a player.

Even though calling the winning function from the
engine was costly at the beginning of the project,
I managed to drastically reduce this time by
improving the performances.

\section{Heuristics}
%TODO add citation to wikipedia for heuristics
%TODO add importance of the heuristic in minmax
Five heuristics were developed in this project
even though only two were made for competitive
use. One heuristic is the naive function for the
second deliverable. Another is a random heuristic
that systematicaly return a new random number. It
was used to spot runtime errors in \textit{MinMax} and
\textit{MonteCarlo} algorithms. The \textit{Basic}
heuristics only check for winning situation and
was used for basic controls.
The two other heuristics are described in
subsection of their own.

@@

\clearpage
\subsection{Vspace}
\label{sec:vspace}
The goal of the Vspace heuristic is to give value to empty
cells in addition to counting rows of similar
symbols. Similarly to the previous heuristic,
Vspace count the size of the lines, but for every
line connected to an empty cell, it uses the
size of this line as a value for the empty cell.

@@

				\draw [->,thick] (A) -- (B);
		\end{tikzpicture}
		\caption{Example of the Vspace process for the
		color player based on the board in
		Figure~\ref{fig:board-example}. the grey cells
		represent cells that contain nothing
		interesting. On the right board, green cells
		are spaces available for the next turn and
		purple cells are spaces that need at least two
		cards to be filled.}
		\label{fig:color-vspace}
\end{figure}
\clearpage
\section{Difficulties}
\subsection{Slowness}
The biggest difficulty encountered during this
project was the slowness of python. In the early
stage of the project, the code was
too ineficient and could not check more than one
move in advance within the limit of three seconds.

The \textit{cProfile} python module is a module
that helps to profile python code. In particular,
it give the amount of time spent in each function
called. This profiler was used to locate the most
time consumming function, then I rely on this information to
restructure the code into better object structure.

However, the result were hardly perceivable and
the code was still too slow for competetive use.
Therefore, I turn to two scientific python
libraries: Numpy and Scipy. Using functions from
these libraries slightly improved the performance
because they were built-in functions. In python,
built-in functions are C functions compiled into a
python module and called from python script.

Finally, to reach a decent executing time, 
one more step was taken and a python module writen
in C was developed.
Figure~\ref{fig:performance-magic} show the
evolution of the performances as the most time
consuming function were adapted to call the
module. However, the python code remains and it is
executed by default. 
Figure~\ref{fig:performance-magic} shows that the
MinMax algorithm is able to check four moves in
advance after half of the key part used the
module.

\begin{figure}
		\includegraphics[width=\textwidth]{perfromances.png}
		\caption{The amount of time (in second) taken
		to run the MinMax algorithm on a board with the
		card "5 B 1" already played. The MinMax
		algorithm checked up to 4 moves in advance and
		it used the Vspace heuristic described in
		Section~\ref{sec:vspace}. The left column
		shows the full python execution time (more
		than 4 minutes) and the other columns show the
		execution time as part of the code start using
		the C module. The third column from the left,
		"checking a move", indicate the time taken
		when the heuristic function and the function
		that check the moves use both the C module.}
		\label{fig:performance-magic}
\end{figure}

\begin{itemize}
		\item Version changing between Python 3.7.2 and 3.5.1 (serialization of regex)
\end{itemize}
\section{Analyze}
\begin{itemize}
		\item How is the pruning depth 3 with a
				recycling moves: no prun = 26s, prun=0.3s.
		\item How are the performance at the tournament.
\end{itemize}
